Depletion capabilities in the OpenMC Monte Carlo particle transport code

### 2020-09-29 manuscript.tex
Fix sentence in conclusions
Fix typos in acknowledgments
Add middle initial for Colin
Address reviewer 1 comments

\author[lanl]{Colin Josey}
\ead{cjosey@lanl.gov}

@@

Monte Carlo (MC) particle transport simulations are now widely used in reactor
design and analysis in a variety of roles thanks to continued advances in
computing. The earliest efforts to couple MC particle transport with depletion
date back to the 1990s~\citep{moore1995inel,trellue1998lanl}. These early
efforts involved taking existing MC and depletion codes, such as
MCNP~\citep{goorley2012nt} and ORIGEN~\citep{croff1983nt}, and putting a
``wrapper'' around them that orchestrated the transfer of solutions from one
code to the other. Thus, the actual transport/depletion codes themselves
required no source code changes. In the past decade, many MC code developers
have chosen to directly include a depletion solver as a first-rate feature,
including MCNPX~\citep{waters2007aip} (which has since been merged into
MCNP6~\citep{goorley2012nt}), Serpent~\citep{leppanen2015ane},
MC21~\citep{griesheimer2015ane}, and RMC~\citep{wang2015ane}.

@@

The equation that governs the transmutation of nuclides by nuclear reactions and
radioactive decay inside an irradiated environment can be written as
\begin{equation}
  \label{eq:depletion}
  \begin{split}
  \frac{dN_i(t)}{dt} = &\sum\limits_j \left [ \int_0^\infty dE \;
  f_{j \rightarrow i}(E) \sigma_j (E, t) \phi(E,t) + \lambda_{j\rightarrow i}
  \right ] N_j(t) \\ &- \left [\int_0^\infty dE \; \sigma_i (E,t) \phi(E,t) +
  \sum\limits_j \lambda_{i\rightarrow j} \right ] N_i(t), \quad i=1,\dots,n ,
  \end{split}
\end{equation}
where
\begin{equation*}
  \begin{split}
  N_i(t) &\equiv \text{density of nuclide $i$ at time $t$} \\
  \sigma_i(E,t) &\equiv \text{transmutation cross section for nuclide $i$ at energy $E$ and time $t$} \\
  \phi(E,t) &\equiv \text{neutron flux at energy $E$ and time $t$} \\
  f_{j \rightarrow i}(E) &\equiv \text{fraction of transmutation reactions in nuclide $j$ that produce nuclide $i$} \\
  \lambda_{j \rightarrow i} &\equiv \text{decay constant for decay modes in nuclide $j$ that produce nuclide $i$} \\
  n &\equiv \text{total number of nuclides.}
  \end{split}
\end{equation*}
Note that we have not included the spatial dependence of the flux or cross
sections. For \cref{eq:depletion} to be accurate, the spatial variation of the
reaction rates over the region that is being depleted should be small. For a
reactor model, this may entail using a unique material in each fuel pin that is
depleted separately.\footnote{OpenMC includes functions for automatically
differentiating materials in cells that are repeated throughout a geometry.}
Ultimately, the user is responsible for deciding what level of spatial
discretization is adequate for the problem at hand.

@@

\subsubsection{Predictor Method}

The simplest numerical method for solving \cref{eq:depletion-matrix} over a
timestep is known as the \emph{predictor} method. It assumes that
$\vect{A}$ is constant over the timestep and assumes the value at the
beginning of the step. The solution then follows directly from
\cref{eq:constant-A}:
\begin{equation}
  \label{eq:predictor}
  \vect{n}_{i+1} = \exp\left(h\vect{A}(\vect{n}_i) \right) \vect{n}_i.
\end{equation}
The end-of-step composition is thus ``predicted'' by the reaction rates at the
beginning of the step. This method has the virtue of requiring a single
transport solution and a single matrix exponential per timestep but is only
first-order accurate.

@@

A series of \emph{predictor--corrector} methods that use multiple stages offer
improved accuracy over the predictor method. The simplest of these methods is
the CE/CM (constant extrapolation, constant midpoint) algorithm, which is the
default algorithm used in MCNP6~\citep{fensin2006tans}. In this method,
$\vect{A}$ is evaluated at the beginning of the step and used to deplete the
composition over half the step. Then, $\vect{A}$ is reevaluated by using the
predicted middle-of-step composition and is used to deplete over the full step.
The algorithm can be expressed as
\begin{equation}
  \begin{split}
    \vect{n}_{i+1/2} &= \exp \left (\frac{h}{2}\vect{A}(\vect{n}_i) \right) \vect{n}_i \\
    \vect{n}_{i+1} &= \exp \left(h \vect{A}(\vect{n}_{i+1/2}) \right) \vect{n}_i.
  \end{split}
\end{equation}
This method requires two transport solutions and two matrix exponential solves
per timestep (twice as expensive as the predictor method) and is second-order
accurate.

@@

All of the integration methods discussed thus far are explicit methods that are
only conditionally stable. In many well-documented cases oscillatory behavior
and diverging solutions have been observed. To address this situation,
\citet{dufek2013ane} proposed an implicit-Euler-like method that is
unconditionally stable. The main idea is to replace $\vect{A}(\vect{n}_i)$ with
$\vect{A}(\vect{n}_{i+1})$ in \cref{eq:predictor}:
\begin{equation}
  \label{eq:implicit}
  \vect{n}_{i+1} = \exp \left( h\vect{A}(\vect{n}_{i+1}) \right) \vect{n}_i.
\end{equation}
This results in a nonlinear equation for $\vect{n}_{i+1}$ that cannot be solved
directly. It is possible, however, to solve for $\vect{n}_{i+1}$ by using a
fixed-point iteration. \citet{dufek2013ane} utilized a stochastic approximation
based on the Robbins--Monro algorithm~\citep{robbins1951ams} in order to
guarantee convergence of the fixed-point iteration while accounting for
stochasticity in the MC transport solver.

The major downside of an algorithm based on \cref{eq:implicit} is that, like the
predictor method, it is only first-order accurate. Implicit methods have also
been applied to higher-order
algorithms~\citep{kotlyar2014ane,kotlyar2016ane,cosgrove2020ane}, thereby
obtaining methods that are unconditionally stable and accurate with longer
timesteps. In OpenMC, stochastic implicit versions of the CE/LI and LE/QI
methods have also been implemented. The basis for the algorithm is to iterate on
$\vect{n}_{i+1}^p$ used in \cref{eq:linear-interp}, obtaining progressively
better estimates that are then used to solve \cref{eq:linear-ode}.

@@

OpenMC does not rely on a single time integration method but rather has several
classes that implement different algorithms. Generally, there is a tradeoff
between the accuracy of the method, the computational expense, and the memory
requirements. When an MC code is used, the expense is dominated by the time to
compute the transport solution, in other words, to evaluate $\vect{A}$ for a
given $\vect{n}$. Thus, the cost of a method scales with the number of
$\vect{A}$ evaluations that are performed per timestep. On the other hand,
methods that require more evaluations generally achieve higher accuracy. The
predictor method requires only one evaluation, and its error converges as
$\mathcal{O}(h)$. The CE/CM method requires two evaluations and is thus twice as
expensive as the predictor method, but achieves an error of $\mathcal{O}(h^2)$.

@@

The PWR pincell problem is a single 2.4\% enriched PWR pincell from the BEAVRS
benchmark~\citep{horelik2013mc,horelik2018mit}. Material compositions and
dimensions from the BEAVRS benchmark were used without modification. The pincell
extends to positive/negative infinity in the $z$ direction and has reflective
boundary conditions in the $x$ and $y$ directions. A linear heating rate of
\SI{174}{\watt\per\centi\meter} was used for power normalization. The pin was
depleted to 50 MWd/kg using the following burnup schedule:
\begin{itemize}
  \item A single 0.1 MWd/kg, 0.4 MWd/kg, and 0.5 MWd/kg timesteps
  \item 1.0 MWd/kg timesteps out to a cumulative burnup of 10.0 MWd/kg
  \item 2.5 MWd/kg timesteps out to a cumulative burnup of 50.0 MWd/kg
\end{itemize}
The short timesteps at the beginning are used to accurately capture the buildup
of xenon isotopes. In both codes, the predictor method was used for time
integration. While the predictor method is the least accurate time integration
method, the purpose of the comparison is to confirm that OpenMC and Serpent
produce comparable results irrespective of the absolute values.

@@

\Cref{fig:pwr-actinides-full} shows the relative difference in predicted
actinide concentrations between OpenMC and Serpent at four burnup values. The
fission products shown in \cref{fig:pwr-actinides-full} have been chosen because
they are either strong neutron absorbers or significant contributors to decay
heat and/or radiotoxicity. We see that most of the differences are less than
0.1\%, and all are within 0.15\%. The differences in $^{235}$U and $^{239}$Pu
concentrations increase slightly with increasing burnup, which most likely
accounts for the 20 pcm difference observed for the full chain in
\cref{fig:pwr-keff}. While not shown, the predicted actinide concentrations for
the reduced chain are nearly identical to those observed for the full chain in
\cref{fig:pwr-actinides-full}. \Cref{fig:pwr-actinides-casl} shows the
differences in actinide concentrations using the CASL depletion chain. Because
the CASL depletion chain differs from the full depletion chain primarily with
respect to how fission products are treated, the differences reported in
\cref{fig:pwr-actinides-casl} closely mirror those in
\cref{fig:pwr-actinides-full}.
\begin{figure}[H]
  \centering
  \includegraphics[width=4in]{figures/pwr_actinides_full.pdf}
  \caption{Relative difference in actinide concentrations between OpenMC and
  Serpent for the PWR pincell model at 5, 12.5, 30, and 50 MWd/kg using the full
  depletion chain.}
  \label{fig:pwr-actinides-full}
\end{figure}
\begin{figure}[H]
  \centering
  \includegraphics[width=4in]{figures/pwr_actinides_casl.pdf}
  \caption{Relative difference in actinide concentrations between OpenMC and
  Serpent for the PWR pincell model at 5, 12.5, 30, and 50 MWd/kg using the CASL
  depletion chain.}
  \label{fig:pwr-actinides-casl}
\end{figure}

The predicted concentration of fission products depends on both the depletion
chain that is used and the method for handling the energy dependence of the FPY.
We begin with results for the full depletion chain and interpolated FPYs, which
most closely match the method used in Serpent; \cref{fig:pwr-fp-full-average}
shows the relative difference in fission product concentrations between OpenMC
and Serpent for this case. Most of the differences are less than 0.2\%, and only
a few fission products show differences up to 0.5\%.
\begin{figure}[H]
  \centering
  \includegraphics[width=4in]{figures/pwr_fp_full_average.pdf}
  \caption{Relative difference in fission product concentrations between OpenMC
  and Serpent for the PWR pincell model at 5, 12.5, 30, and 50 MWd/kg using the
  full depletion chain and interpolated FPYs in OpenMC.}
  \label{fig:pwr-fp-full-average}
\end{figure}

@@

The SFR assembly model is based on an OECD/NEA benchmark for SFR
analysis~\citep{nsc2015}. Specifically, a 2D model of a driver assembly for the
1000 MWth metallic core (MET-1000) was used. The fuel composition corresponded
to the lowest axial region (the 17.16 cm column in Table 2.23
from~\citep{nsc2015}). All dimensions and material compositions can be found in
the original OECD/NEA report. The assembly model extends to positive/negative
infinity in the $z$ direction and has periodic boundary conditions on the outer
hexagonal surfaces. A linear heating rate of \SI{23.3}{\kilo\watt\per\meter} per
pin~\citep{cahalan2007anl} was used for power normalization. The assembly was
depleted for 360 days in 20-day timesteps. As in the PWR pincell case, the
predictor method was used in both codes for time integration.

@@

A new depletion solver has been implemented in the OpenMC Monte Carlo particle
transport code. The depletion solver is implemented primarily in Python 3 and
interfaces with OpenMC's transport solver through Python bindings to an
underlying C++ API, which enables data transfers to happen in memory rather than
through filesystem I/O. The design of the depletion solver relies on a set of
abstract classes that provide modularity and separation between different
components, including the time integration methods, the matrix exponential
solver, and the transport operator. In principle, this modularity would allow
the depletion system to be used with another transport code. Rather than relying
on a single time integration method, OpenMC implements several time integration
schemes, including the predictor method, standard predictor--corrector methods,
and stochastic implicit Euler integrators.
methods~\citep{josey2017phd, josey2017jcp}. To evaluate the matrix exponential,
OpenMC relies on the incomplete partial fraction form of CRAM by default.

@@

This research was supported by the Exascale Computing Project (17-SC-20-SC), a
collaborative effort of the U.S. Department of Energy Office of Science and the
National Nuclear Security Administration. The submitted manuscript was also
suported in part by the U.S. Department of Energy, Office of Science, under
contrct DE-AC02-06CH11357. We gratefully acknowledge the computing resources
provided on Bebop, a high-performance computing cluster operated by the
Laboratory Computing Resource Center at Argonne National Laboratory. This
research made use of the resources of the High Performance Computing Center at
Idaho National Laboratory, which is supported by the Office of Nuclear Energy of
the U.S. Department of Energy and the Nuclear Science User Facilities under
Contract No. DE-AC07-05ID14517.

>>>>

\author[lanl]{Colin J. Josey}
\ead{cjosey@lanl.gov}

@@

Monte Carlo (MC) particle transport simulations are now widely used in reactor
design and analysis in a variety of roles thanks to continued advances in
computing. The earliest efforts to couple MC particle transport with depletion
date back to the 1990s~\citep{moore1995inel,trellue1998lanl}. These early
efforts involved taking existing MC and depletion codes, such as
MCNP~\citep{goorley2012nt} and ORIGEN~\citep{croff1983nt}, and putting a
``wrapper'' around them that orchestrated the transfer of solutions from one
code to the other. Thus, the actual transport/depletion codes themselves
required no source code changes. In the past decade, many MC code developers
have chosen to directly include a depletion solver as a first-rate feature,
including MCNPX~\citep{waters2007aip} (which has since been merged into
MCNP6~\citep{goorley2012nt}), Serpent~\citep{leppanen2015ane},
MC21~\citep{griesheimer2015ane}, RMC~\citep{wang2015ane}, and
Shift~\citep{davidson2016ane}.

@@

The equation that governs the transmutation of nuclides by nuclear reactions and
radioactive decay inside an irradiated environment can be written as
\begin{equation}
  \label{eq:depletion}
  \begin{split}
  \frac{dN_i(t)}{dt} = &\sum\limits_{j=1}^n \left [ \int_0^\infty dE \;
  f_{j \rightarrow i}(E) \sigma_j (E, t) \phi(E,t) + \lambda_{j\rightarrow i}
  \right ] N_j(t) \\ &- \left [\int_0^\infty dE \; \sigma_i (E,t) \phi(E,t) +
  \sum\limits_{j=1}^n \lambda_{i\rightarrow j} \right ] N_i(t), \quad i=1,\dots,n ,
  \end{split}
\end{equation}
where
\begin{equation*}
  \begin{split}
  N_i(t) &\equiv \text{density of nuclide $i$ at time $t$} \\
  \sigma_i(E,t) &\equiv \text{transmutation cross section for nuclide $i$ at energy $E$ and time $t$} \\
  \phi(E,t) &\equiv \text{neutron flux at energy $E$ and time $t$} \\
  f_{j \rightarrow i}(E) &\equiv \text{fraction of transmutation reactions in nuclide $j$ that produce nuclide $i$} \\
  \lambda_{j \rightarrow i} &\equiv \text{decay constant for decay modes in nuclide $j$ that produce nuclide $i$} \\
  n &\equiv \text{total number of nuclides.}
  \end{split}
\end{equation*}
By definition, $f_{i\rightarrow i}$ and $\lambda_{i\rightarrow i}$ are zero so
the summations in \cref{eq:depletion} are effectively over all nuclides except
$i$. Note that we have not included the spatial dependence of the flux or cross
sections. For \cref{eq:depletion} to be accurate, the spatial variation of the
reaction rates over the region that is being depleted should be small. For a
reactor model, this may entail using a unique material in each fuel pin that is
depleted separately.\footnote{OpenMC includes functions for automatically
differentiating materials in cells that are repeated throughout a geometry.}
Ultimately, the user is responsible for deciding what level of spatial
discretization is adequate for the problem at hand.

@@

\subsubsection{Constant Extrapolation Method}

The simplest numerical method for solving \cref{eq:depletion-matrix} over a
timestep is known as the \emph{constant extrapolation} (CE) method\footnote{In
the literature, this algorithm is also sometimes called the fully-explicit
method or the predictor method.}. It assumes that $\vect{A}$ is constant over
the timestep and assumes the value at the beginning of the step. The solution
then follows directly from \cref{eq:constant-A}:
\begin{equation}
  \label{eq:constant_extrapolation}
  \vect{n}_{i+1} = \exp\left(h\vect{A}(\vect{n}_i) \right) \vect{n}_i.
\end{equation}
The end-of-step composition is thus ``predicted'' by the reaction rates at the
beginning of the step. This method has the virtue of requiring a single
transport solution and a single matrix exponential per timestep but is only
first-order accurate.

@@

A series of \emph{predictor--corrector} methods that use multiple stages offer
improved accuracy over the CE method. The simplest of these
methods is the CE/CM (constant extrapolation, constant midpoint) algorithm,
which is the default algorithm used in MCNP6~\citep{fensin2006tans}. In this
method, $\vect{A}$ is evaluated at the beginning of the step and used to deplete
the composition over half the step. Then, $\vect{A}$ is reevaluated by using the
predicted middle-of-step composition and is used to deplete over the full step.
The algorithm can be expressed as
\begin{equation}
  \begin{split}
    \vect{n}_{i+1/2} &= \exp \left (\frac{h}{2}\vect{A}(\vect{n}_i) \right) \vect{n}_i \\
    \vect{n}_{i+1} &= \exp \left(h \vect{A}(\vect{n}_{i+1/2}) \right) \vect{n}_i.
  \end{split}
\end{equation}
This method requires two transport solutions and two matrix exponential solves
per timestep (twice as expensive as the CE method) and is second-order
accurate.

@@

All of the integration methods discussed thus far are explicit methods that are
only conditionally stable. In many well-documented cases oscillatory behavior
and diverging solutions have been observed. To address this situation,
\citet{dufek2013ane} proposed an implicit-Euler-like method that is
unconditionally stable. The main idea is to replace $\vect{A}(\vect{n}_i)$ with
$\vect{A}(\vect{n}_{i+1})$ in \cref{eq:constant_extrapolation}:
\begin{equation}
  \label{eq:implicit}
  \vect{n}_{i+1} = \exp \left( h\vect{A}(\vect{n}_{i+1}) \right) \vect{n}_i.
\end{equation}
This results in a nonlinear equation for $\vect{n}_{i+1}$ that cannot be solved
directly. It is possible, however, to solve for $\vect{n}_{i+1}$ by using a
fixed-point iteration. \citet{dufek2013ane} utilized a stochastic approximation
based on the Robbins--Monro algorithm~\citep{robbins1951ams} in order to
guarantee convergence of the fixed-point iteration while accounting for
stochasticity in the MC transport solver.

The major downside of an algorithm based on \cref{eq:implicit} is that, like the
CE method, it is only first-order accurate. Implicit methods have also been
applied to higher-order
algorithms~\citep{kotlyar2014ane,kotlyar2016ane,cosgrove2020ane}, thereby
obtaining methods that are unconditionally stable and accurate with longer
timesteps. In OpenMC, stochastic implicit versions of the CE/LI and LE/QI
methods have also been implemented. The basis for the algorithm is to iterate on
$\vect{n}_{i+1}^p$ used in \cref{eq:linear-interp}, obtaining progressively
better estimates that are then used to solve \cref{eq:linear-ode}.

@@

OpenMC does not rely on a single time integration method but rather has several
classes that implement different algorithms. Generally, there is a tradeoff
between the accuracy of the method, the computational expense, and the memory
requirements. When an MC code is used, the expense is dominated by the time to
compute the transport solution, in other words, to evaluate $\vect{A}$ for a
given $\vect{n}$. Thus, the cost of a method scales with the number of
$\vect{A}$ evaluations that are performed per timestep. On the other hand,
methods that require more evaluations generally achieve higher accuracy. The CE
method requires only one evaluation, and its error converges as
$\mathcal{O}(h)$. The CE/CM method requires two evaluations and is thus twice as
expensive as the CE method, but achieves an error of $\mathcal{O}(h^2)$.

@@

The PWR pincell problem is a single 2.4\% enriched PWR pincell from the BEAVRS
benchmark~\citep{horelik2013mc,horelik2018mit}. Material compositions and
dimensions from the BEAVRS benchmark were used without modification. The pincell
extends to positive/negative infinity in the $z$ direction and has reflective
boundary conditions in the $x$ and $y$ directions. A linear heating rate of
\SI{174}{\watt\per\centi\meter} was used for power normalization. The pin was
depleted to 50 MWd/kg using the following burnup schedule:
\begin{itemize}
  \item A single 0.1 MWd/kg, 0.4 MWd/kg, and 0.5 MWd/kg timesteps
  \item 1.0 MWd/kg timesteps out to a cumulative burnup of 10.0 MWd/kg
  \item 2.5 MWd/kg timesteps out to a cumulative burnup of 50.0 MWd/kg
\end{itemize}
The short timesteps at the beginning are used to accurately capture the buildup
of xenon isotopes. In both codes, the CE method was used for time
integration. While the CE method is the least accurate time integration
method, the purpose of the comparison is to confirm that OpenMC and Serpent
produce comparable results irrespective of the absolute values.

@@

\Cref{fig:pwr-actinides-full} shows the relative difference in predicted
actinide concentrations between OpenMC and Serpent at four burnup values using
the full depletion chain. We see that most of the differences are less than
0.1\%, and all are within 0.15\%. The differences in $^{235}$U and $^{239}$Pu
concentrations increase slightly with increasing burnup, which most likely
accounts for the 20 pcm difference observed for the full chain in
\cref{fig:pwr-keff}. While not shown, the predicted actinide concentrations for
the reduced chain are nearly identical to those observed for the full chain in
\cref{fig:pwr-actinides-full}. \Cref{fig:pwr-actinides-casl} shows the
differences in actinide concentrations using the CASL depletion chain. Because
the CASL depletion chain differs from the full depletion chain primarily with
respect to how fission products are treated, the differences reported in
\cref{fig:pwr-actinides-casl} closely mirror those in
\cref{fig:pwr-actinides-full}.
\begin{figure}[H]
  \centering
  \includegraphics[width=4in]{figures/pwr_actinides_full.pdf}
  \caption{Relative difference in actinide concentrations between OpenMC and
  Serpent for the PWR pincell model at 5, 12.5, 30, and 50 MWd/kg using the full
  depletion chain.}
  \label{fig:pwr-actinides-full}
\end{figure}
\begin{figure}[H]
  \centering
  \includegraphics[width=4in]{figures/pwr_actinides_casl.pdf}
  \caption{Relative difference in actinide concentrations between OpenMC and
  Serpent for the PWR pincell model at 5, 12.5, 30, and 50 MWd/kg using the CASL
  depletion chain.}
  \label{fig:pwr-actinides-casl}
\end{figure}

The predicted concentration of fission products depends on both the depletion
chain that is used and the method for handling the energy dependence of the FPY.
We begin with results for the full depletion chain and interpolated FPYs, which
most closely match the method used in Serpent; \cref{fig:pwr-fp-full-average}
shows the relative difference in fission product concentrations between OpenMC
and Serpent for this case. The fission products shown in
\cref{fig:pwr-fp-full-average} have been chosen because they are either strong
neutron absorbers or significant contributors to decay heat and/or
radiotoxicity. Most of the differences are less than 0.2\%, and only a few
fission products show differences up to 0.5\%.
\begin{figure}[H]
  \centering
  \includegraphics[width=4in]{figures/pwr_fp_full_average.pdf}
  \caption{Relative difference in fission product concentrations between OpenMC
  and Serpent for the PWR pincell model at 5, 12.5, 30, and 50 MWd/kg using the
  full depletion chain and interpolated FPYs in OpenMC.}
  \label{fig:pwr-fp-full-average}
\end{figure}

@@

The SFR assembly model is based on an OECD/NEA benchmark for SFR
analysis~\citep{nsc2015}. Specifically, a 2D model of a driver assembly for the
1000 MWth metallic core (MET-1000) was used. The fuel composition corresponded
to the lowest axial region (the 17.16 cm column in Table 2.23
from~\citep{nsc2015}). All dimensions and material compositions can be found in
the original OECD/NEA report. The assembly model extends to positive/negative
infinity in the $z$ direction and has periodic boundary conditions on the outer
hexagonal surfaces. A linear heating rate of \SI{23.3}{\kilo\watt\per\meter} per
pin~\citep{cahalan2007anl} was used for power normalization. The assembly was
depleted for 360 days in 20-day timesteps. As in the PWR pincell case, the CE
method was used in both codes for time integration.

@@

A new depletion solver has been implemented in the OpenMC Monte Carlo particle
transport code. The depletion solver is implemented primarily in Python 3 and
interfaces with OpenMC's transport solver through Python bindings to an
underlying C++ API, which enables data transfers to happen in memory rather than
through filesystem I/O. The design of the depletion solver relies on a set of
abstract classes that provide modularity and separation between different
components, including the time integration methods, the matrix exponential
solver, and the transport operator. In principle, this modularity would allow
the depletion system to be used with another transport code. Rather than relying
on a single time integration method, OpenMC implements several time integration
schemes, including the constant extrapolation method, standard
predictor--corrector methods, and stochastic implicit Euler integrators. To
evaluate the matrix exponential, OpenMC relies on the incomplete partial
fraction form of CRAM by default.

@@

This research was supported by the Exascale Computing Project (17-SC-20-SC), a
collaborative effort of the U.S. Department of Energy Office of Science and the
National Nuclear Security Administration. The submitted manuscript was also
supported in part by the U.S. Department of Energy, Office of Science, under
contract No. \mbox{DE-AC02-06CH11357}. We gratefully acknowledge the computing
resources provided on Bebop, a high-performance computing cluster operated by
the Laboratory Computing Resource Center at Argonne National Laboratory. This
research made use of the resources of the High Performance Computing Center at
Idaho National Laboratory, which is supported by the Office of Nuclear Energy of
the U.S. Department of Energy and the Nuclear Science User Facilities under
Contract No. \mbox{DE-AC07-05ID14517}.
